---
title: "Spring 2021 Waste Audit"
author: ""
subtitle: Duke Student Government Services and Sustainability Committee
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
library(tinytex)
```

```{r upload-and-clean-data}
waste_audit <- read.csv(file = 'waste_audit_datasheet.csv') %>% 
  select(-c(X, X26)) %>% 
  slice(1:22) %>% 
  select(-prop_pieces_trash, -prop_pieces_recycle, -prop_pieces_compost, -prop_trash_weight, -prop_recycle_weight, -prop_compost_weight) %>% 
  mutate(prop_pieces_trash = pieces_trash / total_piece_volume,
         prop_pieces_recycle = pieces_recycle / total_piece_volume,
         prop_pieces_compost = pieces_compost / total_piece_volume,
         prop_trash_weight = weight_trash / total_can_weight,
         prop_recycle_weight = weight_recycle / total_can_weight,
         prop_compost_weight = weight_compost / total_can_weight,
         weight_recycle_lbs = weight_recycle,
         weight_trash_lbs = weight_trash,
         weight_compost_lbs = weight_compost)
```

## Content Comparisons
- average weight of a piece of trash, composting, and recycling item
```{r average-piece-weight}
waste_audit %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash_lbs), 
            total_recycle_weight = sum(weight_recycle_lbs),
            total_compost_weight = sum(weight_compost_lbs),
            average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

- comparative starting volume of pieces/weight between trash, recycle, and
compost streams
```{r starting-volumes}
waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(avg_starting_weight = mean(total_can_weight),
            avg_starting_pieces = mean(total_piece_volume))
```

## Quantifying Missorted Waste
```{r missorted}
waste_audit <- waste_audit %>% 
  mutate(perc_missorted_c = case_when(original_stream == "Trash" ~ (weight_recycle_lbs + weight_compost_lbs) / total_can_weight, original_stream == "Recycling" ~ (weight_trash_lbs + weight_compost_lbs) / total_can_weight),
         missorted_count = case_when(original_stream == "Trash" ~ (weight_recycle_lbs + weight_compost_lbs), original_stream == "Recycling" ~ (weight_trash_lbs + weight_compost_lbs)),
         sorted_count = total_can_weight - missorted_count)
```

- percent of total waste that is missorted

```{r missorted-total-waste}
pie_total <- waste_audit %>% 
  summarize(sorted = sum(sorted_count),
            missorted = sum(missorted_count))
pie_total
slices <- c(31, 137.2)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Overall Contamination for All Streams")
# source: https://www.statmethods.net/graphs/pie.html
```

- Bootstrap Sample OVERALL WASTE

bootstrap sample

```{r bootstrap-sample-with-compost}
set.seed(201)
boot_missorted <- waste_audit %>% 
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-with-compost}
ci_boot_missorted <- get_ci(boot_missorted, level = 0.95)
ci_boot_missorted
```

bootstrap distribution

```{r bootstrap-vis-with-compost}
visualize(boot_missorted) +
  shade_ci(ci_boot_missorted) +
  labs(title = "Simulated Bootstrap Distribution of Overall Missorted Waste", x = "Sample Mean Missorted Overall Waste", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-with-compost, eval = FALSE}
mean(waste_audit$perc_missorted_c) + c(-1, 1) * 2.05 * sd(waste_audit$perc_missorted_c) / sqrt(22)
```

tdist: 0.6798681, 0.8236861

- percent of trash that is missorted

```{r missorted-trash}
pie_trash <- waste_audit %>% 
  mutate(missorted = weight_recycle_lbs + weight_compost_lbs) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(weight_trash_lbs),
            missorted = sum(missorted)) %>%
  slice(2)
pie_trash
slices <- c(18.8, 114.6)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Trash")
```

- Bootstrap Sample TRASH ONLY

bootstrap sample
```{r bootstrap-sample-trash}
boot_table_trash <- waste_audit %>%
  slice(1:11, 21:22)
set.seed(201)
boot_missorted_trash <- boot_table_trash %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-trash}
ci_boot_missorted_trash <- get_ci(boot_missorted_trash, level = 0.95)
ci_boot_missorted_trash
```

bootstrap distribution

```{r bootstrap-vis-trash}
visualize(boot_missorted_trash) +
  shade_ci(ci_boot_missorted_trash) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Trash", x = "Sample Mean Missorted Trash", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-trash, eval = FALSE}
mean(boot_table_trash$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_trash$perc_missorted_c) / sqrt(13)
```

tdist: 0.7987486, 0.8953845

- percent of recycling that is missorted
    
```{r missorted-recycle}
pie_recycle <- waste_audit %>% 
  mutate(missorted = weight_trash_lbs + weight_compost_lbs) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(weight_recycle_lbs),
            missorted = sum(missorted)) %>%
  slice(1)
pie_recycle
slices <- c(12.2, 22.6)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Recycle")
```

- Bootstrap Sample RECYCLING ONLY

bootstrap sample

```{r bootstrap-sample-recycling}
boot_table_recycling <- waste_audit %>%
  slice(12:20)
set.seed(201)
boot_missorted_recycling <- boot_table_recycling %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-recycling}
ci_boot_missorted_recycling <- get_ci(boot_missorted_recycling, level = 0.95)
ci_boot_missorted_recycling
```

bootstrap distribution

```{r bootstrap-vis-recycling}
visualize(boot_missorted_recycling) +
  shade_ci(ci_boot_missorted_recycling) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Recycling", x = "Sample Mean Missorted Recycling", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-recycling, eval = FALSE}
mean(boot_table_recycling$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_recycling$perc_missorted_c) / sqrt(9)
```

tdist: 0.5079304, 0.7203432
