---
title: "Spring 2021 Waste Audit"
author: ""
subtitle: Duke Student Government Services and Sustainability Committee
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
library(tinytex)
library(sf)
```

```{r upload-and-clean-data}
waste_audit <- read.csv(file = 'waste_audit_datasheet.csv') %>% 
  select(-c(X, X26)) %>% 
  slice(1:22) %>% 
  select(-prop_pieces_trash, -prop_pieces_recycle, -prop_pieces_compost, -prop_trash_weight, -prop_recycle_weight, -prop_compost_weight) %>% 
  mutate(prop_pieces_trash = pieces_trash / total_piece_volume,
         prop_pieces_recycle = pieces_recycle / total_piece_volume,
         prop_pieces_compost = pieces_compost / total_piece_volume,
         prop_trash_weight = weight_trash / total_can_weight,
         prop_recycle_weight = weight_recycle / total_can_weight,
         prop_compost_weight = weight_compost / total_can_weight)
```

# Introduction

hypothesis

goal:

- judge importance of waste education on campus
- identify ways that we could educate students on waste more successfully
- identify ways that the university could improve waste practices
- quantify the amount of compost produced
- calculate how much composting could be diverted from landfills
- judge the value of the waste audit; should it be recreated?
- how did COVID impact waste on campus? how did COVID impact waste numbers on campus?

- percent of students living on campus
- how many students living on campus; can we make an estimate as to how much waste the average meal produces?
- global trash statistics
- sustainable dining goals in implementing compostables during COVID
- composting surveys
- this comes at a time after (or right before???) the university placed composting awareness posters on the doors of WU

action items:

- call a local composting company (or the composting company of a university of similar size) and ascertain how much compost they collect
- contact the administrators of the tray collector resource in WU to see if any data has been collected incrementally or since the implementation of this infrastructure

# Methodology

## Content Comparisons
- average weight of a piece of trash, composting, and recycling item
```{r average-piece-weight}
waste_audit %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash), 
            total_recycle_weight = sum(weight_recycle),
            total_compost_weight = sum(weight_compost)) %>% 
  summarize(average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

- comparative starting volume of pieces/weight between trash and recycle streams
```{r starting-volumes}
waste_audit %>% 
  mutate(original_weight = weight_trash + weight_recycle + weight_compost,
         original_pieces = pieces_trash + pieces_recycle + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(avg_starting_weight = mean(original_weight),
            avg_starting_pieces = mean(original_pieces))
```

```{r bar-plot-original-stream-piece-comparison}
data <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(total_trash = sum(pieces_trash),
            total_recycle = sum(pieces_recycle),
            total_compost = sum(pieces_compost)) %>% 
  select(total_trash, total_recycle, total_compost)

data <- as.matrix(data)

colnames(data) <- c("trash", "recycle", "compost")

rownames(data) <- c("Recycle", "Trash")

color <- c("black", "#3265ab", "#2ca816")

barplot(t(data), beside = FALSE, col = color, ylim = c(0, 2500),
        main = "Missorted Waste", xlab = "Original Stream Designation", ylab = "Total Pieces")
  legend("topleft", colnames(data), pch = 15, col = color, horiz = T, bty = "n")
  
# source: https://stackoverflow.com/questions/61506408/grouped-barchart-in-r-with-4-variables
```

## Quantifying Missorted Waste

```{r missorted}
waste_audit <- waste_audit %>% 
  mutate(perc_missorted_c = case_when(original_stream == "Trash" ~ (pieces_recycle + pieces_compost) / total_piece_volume, original_stream == "Recycling" ~ (pieces_trash + pieces_compost) / total_piece_volume),
         perc_missorted = case_when(original_stream == "Trash" ~ pieces_recycle / (pieces_trash + pieces_recycle), original_stream == "Recycling" ~ pieces_trash / (pieces_trash + pieces_recycle)))
```

```{r bar-plot-original-stream-percent}
makeup <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(average_trash = mean(prop_pieces_trash),
            average_recycle = mean(prop_pieces_recycle),
            average_compost = mean(prop_pieces_compost)) %>% 
  select(average_trash, average_recycle, average_compost)

makeup <- as.matrix(makeup)

colnames(makeup) <- c("trash", "recycle", "compost")

rownames(makeup) <- c("Recycle", "Trash")

color <- c("black", "#3265ab", "#2ca816")

barplot(t(makeup), beside = FALSE, col = color, ylim = c(0, 1.1),
        main = "Missorted Waste", xlab = "Original Stream Designation", ylab = "Percent")
legend("topleft", colnames(makeup), pch = 15, col = color, horiz = T, bty = "n")
```

- percent of trash that is missorted
    
```{r missorted-trash}
pie_trash <- waste_audit %>% 
  mutate(missorted = pieces_recycle + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_trash),
            missorted = sum(missorted)) %>% 
  slice(2)
pie_trash

slices <- c(555, 1639)
lbls <- c("properly sorted", "missorted")

percent <- slices / sum(slices) * 100

percent

pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Trash")

# source: https://www.statmethods.net/graphs/pie.html
```

```{r percent-of-trash-that-was-landfill}
pie_trash_no_compost <- waste_audit %>% 
  mutate(missorted = pieces_recycle) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_trash),
            missorted = sum(missorted)) %>% 
  slice(2)
pie_trash_no_compost

slices <- c(555, 365)
lbls <- c("properly sorted", "missorted")

percent <- slices / sum(slices) * 100

percent

pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Percent of Trash that was Landfill")
```

- percent of recycling that is missorted include compost
    
```{r missorted-recycle}
pie_recycle <- waste_audit %>% 
  mutate(missorted = pieces_trash + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_recycle),
            missorted = sum(missorted)) %>% 
  slice(1)
pie_recycle

slices <- c(174, 489)
lbls <- c("properly sorted", "missorted")

percent <- slices / sum(slices) * 100

percent

pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Recycle")
```

```{r percent-of-recycling-that-was-recyclable-pie}
pie_rec_no_compost <- waste_audit %>% 
  mutate(missorted = pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_recycle),
            missorted = sum(missorted)) %>% 
  slice(1)
pie_rec_no_compost

slices <- c(174, 169)
lbls <- c("properly sorted", "missorted")

percent <- slices / sum(slices) * 100

percent

pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Percent of Recycling that was Recyclable")
```

## Campus Compost
- percent of trash and recycling that could be compost
    
```{r compost-in-trash}
compostable_trash <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(2)
compostable_trash

slices <- c(1274, 920)
lbls <- c("compostable", "noncompostable")

percent <- slices / sum(slices) * 100

percent

pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Trash that is Compostable")
```
    
```{r compost-in-recycling}
compostable_recycle <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(1)
compostable_recycle

slices <- c(320, 343)
lbls <- c("compostable", "noncompostable")

percent <- slices / sum(slices) * 100

percent

pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Recycling that is Compostable")
```

- volume of compostables
    bar graph, value
    
```{r compost-volume-values}
volume_bar <- waste_audit %>% 
  summarize(total_pieces_compost = sum(pieces_compost),
            total_weight_compost = sum(weight_compost),
            total_pieces_audited = sum(total_piece_volume),
            total_weight_audited = sum(total_can_weight),
            perc_pieces_compost = total_pieces_compost / total_pieces_audited * 100,
            perc_weight_compost = total_weight_compost / total_weight_audited * 100)

makeup <- as.matrix(volume_bar)

volume_bar <- volume_bar %>% 
  select(perc_pieces_compost, perc_weight_compost)

colnames(volume_bar) <- c("Pieces", "Weight (lbs)")

color <- c("#2ca816", "#8cd9b3")

barplot(t(volume_bar), beside = TRUE, col = color, ylim = c(0, 100),
        main = "Percent Total Compostable", xlab = "Method of Quantifying Waste", ylab = "Percent")
legend("topleft", colnames(volume_bar), pch = 15, col = color, horiz = T, bty = "n")
```

## Differences in Location
-difference between big belly compactors and blue bins (is this because multiple streams are offered through Big Belly Compactors?)
 cannot be determined with high enough confidence; sample size of blue single sream trash bins has a margin of error of 49% at 95% confidence
 
-location of the can - use simple features (sf) multi-point data and coordinates in dataset; how do I do this?
    spatial data - how to do a writeup on this type of spatial data, especially when out of immediate visual context of what these coordinates mean on campus and in relation to the location of the cans

```{r, eval = FALSE}
coords <- waste_audit %>% 
  select(x_coordinate, y_coordinate) %>% 
  mutate(x_coordinate = as.numeric(x_coordinate),
         y_coordinate = as.numeric(y_coordinate))

multipoint_matrix <- list(rbind(coords, 'x_coordinate', 'y_coordinate'))

st_multipoint(multipoint_matrix, dim = "XY")

#code source: https://geocompr.robinlovelace.net/spatial-class.html but is now butchered
```

- can compare near exit maybe? looks like the rest of the qualitative samples are too small to compare

## Sample Size and Confidence

# Bootstrap Sample that does not include composting in the total piece volume to determine contamination rates of just trash and recycle (excluding the option to compost)

null hypothesis and alternative hypothesis

bootstrap sample

```{r bootstrap-sample}
set.seed(405)

boot_missorted <- waste_audit %>% 
  specify(response = perc_missorted) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval}
ci_boot_missorted <- get_ci(boot_missorted, level = 0.95)
ci_boot_missorted
```

bootstrap distribution

```{r bootstrap-vis}
visualize(boot_missorted) +
  shade_ci(ci_boot_missorted) +
  labs(title = "Simulated Bootstrap Distribution", x = "Sample Mean", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution, eval = FALSE}
mean(waste_audit$perc_missorted) + c(-1, 1) * 2.05 * sd(waste_audit$perc_missorted) / sqrt(22)
```

tdist: 0.4044905, 0.5056520

# Bootstrap Sample that DOES include composting in the total piece volume to determine holistic contamination rates

null hypothesis and alternative hypothesis

bootstrap sample

```{r bootstrap-sample}
set.seed(405)

boot_missorted <- waste_audit %>% 
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval}
ci_boot_missorted <- get_ci(boot_missorted, level = 0.95)
ci_boot_missorted
```

bootstrap distribution

```{r bootstrap-vis}
visualize(boot_missorted) +
  shade_ci(ci_boot_missorted) +
  labs(title = "Simulated Bootstrap Distribution", x = "Sample Mean", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution, eval = FALSE}
mean(waste_audit$perc_missorted_c) + c(-1, 1) * 2.05 * sd(waste_audit$perc_missorted_c) / sqrt(22)
```

tdist: 0.7146358, 0.7902705

# Results

## Content Comparisons

## Quantifying Missorted Waste

## Campus Compost

## Differences in Location

## Sample Size and Confidence

# Discussion

- error: human error in sorting
- error: some contamination spreads (e.g. containers filled with liquid)
- discuss time of day and year that the data was collected
- note: this does not account for all waste; waste disposed of in dorms, WU conveyor belt, BC, library, quad trash bins, gym (Red Mango), East Campus (Marketplace, dorms)
- semi-random sampling (all recycling bins chosen, which dictated our choices in picking from the trash bins
- error: subjectivity in what it means to be proximal to an exit door, stairwell mouth, etc.

# Call to Action



# Cited References

1. 