---
title: "Spring 2021 Waste Audit"
author: "Hailey Ross, Shreya Joshi, Danny Wu, Chaya Brennan Agarwal"
subtitle: Duke Student Government Services and Sustainability Committee
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
```

```{r upload-and-clean-data}
waste_audit <- read.csv(file = 'waste_audit_datasheet.csv') %>% 
  select(-c(X, X26)) %>% 
  slice(1:22)
```

# Introduction

goal:
- judge importance of waste education on campus
- identify ways that we could educate students on waste more successfully
- identify ways that the university could improve waste practices
- quantify the amount of compost produced
- calculate how much composting could be diverted from landfills
- judge the value of the waste audit; should it be recreated?
- how did COVID impact waste on campus? how did COVID impact waste numbers on campus?

- percent of students living on campus
- how many students living on campus; can we make an estimate as to how much waste the average meal produces?
- global trash statistics
- sustainable dining goals in implementing compostables during COVID
- composting surveys
- this comes at a time after (or right before???) the university placed composting awareness posters on the doors of WU

# Methodology

## Content Comparisons
- average weight of a piece of trash, composting, and recycling item
```{r average-piece-weight}
waste_audit %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash), 
            total_recycle_weight = sum(weight_recycle),
            total_compost_weight = sum(weight_compost)) %>% 
  summarize(average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

- comparative starting volume of pieces/weight between trash and recycle streams
    bar graphs
```{r starting-volumes}
waste_audit %>% 
  mutate(original_weight = weight_trash + weight_recycle + weight_compost,
         original_pieces = pieces_trash + pieces_recycle + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(avg_starting_weight = mean(original_weight),
            avg_starting_pieces = mean(original_pieces)) %>% 
  ggplot(data = waste_audit) +
  geom_bar()
  
```
in the middle of relearning ggplot^ oops

## Quantifying Missorted Waste
- percent of trash that is missorted
    pie chart, value, bar chart
- percent of recycling that is missorted
    pie chart, value, bar chart

## Campus Compost
- percent of trash and recycling that could be compost
    value, pie chart, bar graphs (each bar split into three sections)
- volume of compostables
    bar graph, value

## Differences in Location
-difference between big belly compactors and blue bins (is this because multiple streams are offered through Big Belly Compactors?)
    bar graphs, box plots
-location of the can
    spatial data

## Sample Size and Confidence
-sample size; can this waste audit's data conclusively say anything about campus, or can it merely be used to suggest the need for a more extensive audit?
    confidence intervals?

# Results

## Content Comparisons

## Quantifying Missorted Waste

## Campus Compost

## Differences in Location

## Sample Size and Confidence

# Discussion

- error: human error in sorting
- error: some contamination spreads (e.g. containers filled with liquid)
- discuss time of day and year that the data was collected
- note: this does not account for all waste; waste disposed of in dorms, WU conveyor belt, BC, library, quad trash bins, gym (Red Mango), East Campus (Marketplace, dorms)

# Call to Action



# Cited References

1. 